### 2020-09-01 paper_review/response_20200826.txt
update referee response

Dear Dr. Endl,

We thank you for organizing the review of our paper, and also
thank the referee for constructive criticism.  We have considered
the referee's remarks carefully and revised our manuscript
accordingly.

Below, we have reproduced the relevant portions of the referee's
report, alongside our responses. A list of changes is appended to
the manuscript using the trackchanges AASTeX macros. A figure
supporting the verification of our SB2 injection-recovery test
has also been uploaded for inspection.

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
It appears that TOI 837 b was previously noted as a candidate
planet and IC 2602 member in Nardiello et al. 2020
(https://ui.adsabs.harvard.edu/abs/2020MNRAS.495.4924N/abstract).
If the authors were unaware, it is still worth mentioning this in
the text.

RESPONSE
----------
> Thank you for noting the omission. We have added the relevant
> citation in Section 2.1.

******************************************
REFEREE COMMENT
----------
Sec. 2.1
A number of validation tests are listed, but no detail is
provided about how these tests work. For example, what is the
ghost diagnostic test?

RESPONSE
----------
> For the sake of brevity, we prefer to continue with our
> refererence to Twicken et al., 2018. The optical ghost
> diagnostic test in the SPOC pipeline (Twicken et al., 2018,
> Section 3.7) flags threshold crossing events that are likely to
> be caused by reflections of light from bright sources between
> CCDs. Since few people know what it is, and it will therefore
> confuse rather than clarify, we have omitted it from the
> manuscript.

******************************************
REFEREE COMMENT
----------
Sec. 2.1
Have the authors performed a by-eye search for additional
transits in the light curve?

RESPONSE
----------
> We did perform a by-eye search for additional transits in the
> light curve, and we did not find any compelling events. The
> logical next step -- quantifying our sensitivity through
> injection-recovery simulations -- would be an interesting
> exercise.  For it to be valid, care would need to be taken to
> marginalize out dependence on the detrending method (e.g.,
> David et al 2019 and 2020's experience which led to the
> four-planet V1298 Tau system).  Given the required level of
> effort, we feel this exercise to be outside the scope of the
> present analysis.

******************************************
REFEREE COMMENT
----------
Sec. 2.2
Given that Gaia typically resolves binaries at separations > 1
arcsec, is the blending scenario between Star A and TOI 837
plausible?

RESPONSE
----------
> Referring to the "Gaia" line in the bottom sub-panel of our
> Figure 6 (which came from Rizzuto et al., 2018), at 2.3
> arcseconds of separation between Star A and TOI 837, Gaia
> typically resolves binaries to brightness contrasts of ΔG ~=
> 5.0. The actual brightness difference between the two is 4.7
> G-band magnitudes. So yes, we believe that the blending
> scenario is still plausible, primarily because Star A is much
> fainter than TOI 837, placing it near the resolution threshold.

******************************************
REFEREE COMMENT
----------
Sec. 2.2
It seems another line of evidence against Star B (rho=5.4 arcsec)
as the host is the difference imaging centroid test mentioned in
Section 2.1, which they state localizes the transit to within 2
arcseconds of the target.

RESPONSE
----------
> Correct -- we have modified the wording in Sec 2.4 to clarify
> this point.

******************************************
REFEREE COMMENT
----------
Given that TOI 837 and Star A have a large contrast but similar
distances, I would expect the two stars to have a large
difference in their expected densities. Can the mean stellar
density from the transit fit be used to further argue that the
planet orbits TOI 837?

RESPONSE
----------
FIXME FIXME FIXME ... maybe it can...

> -Unfortunately, we do not feel that this argument based on the
> stellar density can be convincingly made. The reason is that
> the transit is grazing, and so our "measurement" of the stellar
> density is in fact largely prior-dominated, rather than being
> determined by the transit data. In other words: in fitting the
> transit data, we assume a stellar radius and logg, and we use
> restrictive priors inferred from the star's position on the
> cluster isochrone (see Table 5). If we were to let those priors
> be much less restrictive (say, with logg being uniform from
> 3.8-4.8, or similar), then we would not be able to break the
> classic degeneracy between the impact parameter and Rp/R*. ...
> (....or would this work regardless?)

******************************************
REFEREE COMMENT
----------
The orientation of Figure 2 is confusing. The axes are labeled
as right ascension and declination, but the lines of constant RA
and Dec are at an angle.

RESPONSE
----------
> We have updated Figure 2, and opted to drop the axis labels
> entirely. We updated the caption to emphasize the orientation
> as given by the compass.

******************************************
REFEREE COMMENT
----------
Sec. 2.3
Suggest swapping the order of Figures 3 & 4, as the high
resolution imaging is discussed before the ground-based light
curves.

RESPONSE
----------
> Done, thank you for the suggestion.

******************************************
REFEREE COMMENT
----------
Sec. 2.5.1
For the secondary spectral line injection, it is not clear
whether a synthetic spectrum was added to the raw spectrum before
deriving the LSD profile - or whether a synthetic LSD profile was
added to the true profile. It is not clear that the two
procedures would yield the same results.

RESPONSE
----------
> We have updated the manuscript to clarify that the SB2
> grid-search injection and recovery exercise was conducted with
> the secondary signal injected into the LSD profiles, not the
> spectra. 
> 
> To test the validity of this technique, we injected a second
> set of lines to the original spectra, and compared the
> resulting derived LSD profiles. To inject a second set of lines
> to the raw spectrum, we made use of a high SNR CHIRON spectrum
> of a G0 RV standard star, scaled its flux, and blended it with
> the averaged observed spectrum of TOI-837.  We then performed
> the LSD routine on the blended spectrum to derive its LSD
> broadening profile as we would for a normal observation.
> 
> The attached figure shows the resulting LSD profiles from our
> injection test. The left column (black) shows the LSD profiles
> when the secondary signals were injected directly into the LSD
> profile. The right column (red) shows the results of our
> blended spectrum analysis. There is minimal difference between
> the two injection techniques.

******************************************
REFEREE COMMENT
----------
Sec. 3.1.3
The outer separation of the "Not SB2" constraint in the top panel
of Fig. 6 seems optimistic. Wouldn't the velocity separation of a
binary with a=100 AU be less than the broadening of 15 km/s?

RESPONSE
----------
> FIXME TODO

******************************************
REFEREE COMMENT
----------
Sec. 3.1.4
It is not clear how the range of K-P parameter space stated was
explored in detail. MCMC sampling is inefficient for multimodal
distributions, and given the sparse sampling of the RV time
series, it seems likely that multiple modes exist. Was K fit
independently over a grid of periods? Orbital inclination would
also be an important parameter to include. If models were used to
transform mass ratio limits to contrast limits this should be
explicitly stated.

RESPONSE
----------
> FIXME TODO

******************************************
REFEREE COMMENT
----------
Sec. 3.2
The following sentence is somewhat imprecise: "The relevant
populations need to be modeled at the Monte Carlo level." Monte
Carlo is a sampling method, whereas the "modeling" step lies in
the choice of probability distributions for the variables being
sampled. Also, is it the sampling that gives a more reliable
estimate of the FPP or is it the Bayesian framework, which
incorporates priors from population occurrence rates?

RESPONSE
----------
> Thank you for this point. We have revised the relevant
> sentences in Sec 3.2 to clarify that the probabilistic
> framework is the important aspect, not the sampling technique.

******************************************
REFEREE COMMENT
----------
Does the secondary eclipse non-detection apply to a specific
phase (e.g. 0.5) or across all phases?

RESPONSE
----------
> All phases -- we have revised the text in Sec 3.2 to clarify
> this point.

******************************************
REFEREE COMMENT
----------
Sec. 4.1.2
The statement that TOI 837 lies along the empirical cluster
isochrone and this limits the presence of a companion to be <50%
the brightness seems to be misstated. I believe the authors mean
that TOI 837 appears to lie below the equal-brightness binary
sequence (which sits 0.75 mag above the single star sequence).
This limits a hypothetical companion to be less than the
brightness of the target star (or less than 50% the total system
brightness).

RESPONSE
----------
> Thank you for raising this point. Our original writing was in
> error, because we meant something slightly stronger. If TOI 837
> were an equal-brightness binary system, it would be 0.75
> magnitudes above the single star sequence. However TOI 837 is
> not only not on the equal-brightness binary sequence -- it is
> not at all elevated with respect to the single-star sequence. A
> binary companion with say half the brightness of the primary
> would make the system 0.44 mags brighter than the single star
> sequence. Based on the intrinsic scatter in the HR diagram,
> this is also ruled out.  We (imprecisely) meant this second
> point, and have modified the text accordingly.

******************************************
REFEREE COMMENT
----------
Sec. 4.2.4
It is not immediately clear that the mass-temperature-age
combination of ~1.2 Msun, 5960 K, and 30-45 Myr is incorrect. The
mass-temperature relation for stars in this mass range is not the
same on the pre-main sequence as it is on the main sequence.

RESPONSE
----------
> Looking at the MIST isochrones, it seems that a
> solar-metallicity, 1.1Msun star reaches the ZAMS around
> log(age) of 7.5, or ~32 Myr. So, TOI 837 seems likely to be
> roughly at the ZAMS, rather than particularly early in its
> pre-main sequence contraction.
>
> Regardless, we have modified the wording here to reflect that
> our preference for "Method 2" is linked to our ability to
> more easily quantify its uncertainties (statistical and
> systematic -- see below).

******************************************
REFEREE COMMENT
----------
Sec. 4.2.4
The authors are encouraged to explore the effect of adopting a
different set of evolutionary models. The Dartmouth models of
Feiden (2016), which include a magnetic field prescription, have
been shown to more accurately reproduce the parameters of
pre-main sequence stars. In the mass range of TOI 837 the effects
may not be dramatic, but are likely to be significant compared to
the quoted uncertainty on e.g. the stellar mass. I would expect
that a 10% uncertainty on the stellar mass would be more
realistic (in the sense that it would accommodate systematic
uncertainties) than the currently quoted 1% uncertainty.

RESPONSE
----------
> Thank you for this suggestion. We have amended Section 4.2.4 to include an
> application of the same isochrone technique, but using the PARSEC models
> rather than MIST. The corresponding differences between MIST and PARSEC are
> now described in the text, and are taken as "systematic uncertainties", and
> propagated through the stellar parameters in Table 4, as well as the transit
> fit.
>
> As a note, we also checked the Feiden (2016) Dartmouth models, but from what
> we could tell, no synthetic Gaia-band photometry was available. Since this
> complicates the comparison between between the Gaia photometry and the
> models, we opted for the PARSEC models.

******************************************
REFEREE COMMENT
----------
Sec. 4.3
The authors perform a joint fit of the TESS and ground-based
photometry to derive a more precise ephemeris, but fit only the
TESS data for the remaining planet parameters. If the errors on
the respective data sets are properly scaled (including jitter
terms if needed), the extra noise from the ground-based
photometry should not significantly impact the inferred
parameters when including those data. This is merely a comment,
and no extra work is necessary.

RESPONSE
----------
> Noted, with thanks.

******************************************
REFEREE COMMENT
----------
Given the planet's young age, it would be interesting to know if
the transit photometry meaningfully constrains the eccentricity
(which along with impact parameter could explain the short
duration). Was a transit fit including eccentricity performed? If
not, there are methods to infer the eccentricity from the
posteriors of a circular orbit fit (see Dawson & Johnson 2012).

RESPONSE
----------
> FIXME TODO

******************************************
REFEREE COMMENT
----------
-"The condition for a grazing transit is whether the impact
parameter b is below 1 - Rp/R*" → below should be replaced with
"above" or "greater than".

RESPONSE
----------
> Corrected, thank you.

******************************************
REFEREE COMMENT
----------
Discussion
More emphasis could be placed on the unusually large size of the
planet. Some speculative discussion on the nature of this
inflation could be warranted. Figure 11 might also be revised to
also reflect the effect of insolation (or the data plotted could
be restricted to close-in planets).

RESPONSE
----------
> Thank you for this suggestion. We have added a sub-plot in
> Figure 11 showing planet sizes and orbital periods, with the
> sub-100 Myr old planets emphasized.  We considered also
> including a plot with planet insolations on the ordinate
> instead of orbital periods, but the NASA Exoplanet Archive
> column for planet insolation is much less complete, presumably
> owing to the sparsity of available stellar luminosities.  We
> have also added some words in the discussion to the effect that
> that the young planets could be large because of i) a selection
> effect or ii) atmospheric escape.

******************************************
REFEREE COMMENT
----------
The paper ends on a note about nodal precession. Is it simple to
compute the nodal precession timescale given some reasonable
assumptions? This might help determine how likely it is that the
effect may be observed in the near term.

RESPONSE
----------
> Our thanks for this comment, which prompted us to calculate the
> precession frequency of the orbital axis due to the stellar
> oblateness using equations from Anderson+2018 and Lai+2018. We
> found that given the observed parameters of the system and an
> assumed stellar Love number (k2) equal to that of the Sun, the
> precession period would be 45,000 years. To complete say 1% of
> a precession cycle, which might roughly be what is needed to
> observe noticeable changes, we would need to wait ~450 years.
>
> We have correspondingly modified the last sentence to say
> something more generic, but actually true: more photometry
> should improve our ability to constrain the orbital
> eccentricity, and will let us search for additional transiting
> planets.

******************************************
ADDITIONAL CHANGES

A few additional changes are noted in the "List of Changes"
appended at the end of the manuscript:

* Citations to Donati+2020 and Damasso+2020 are added to clarify
  that the planetary nature of the RV-detected hot Jupiters in
  Taurus has been debated.